---
output:
  word_document: default
  html_document: default
---

```{r}
library(data.table) # For converting data into time series format
library(ggplot2) # To plot various plots
library(fpp2) # For examining seasonality graphically
library(forecast) # For various functions related to Time series
library(stats) # For applying tests like acf, Ljung-Box Tests
library(tseries)# For applying Dickey Fuller test
library(MLmetrics) # For applying various metrics related to ML
```

```{r}

data = gas

str(data)
start(data)
end(data)
frequency(data)
cycle(data)
```

Reading the data as a time-series object

```{r}
ts.data = ts(gas, start = c(1956,1), end = c(1995,8), frequency = 12)

summary(ts.data)
str(ts.data)

```
The Gas Production values range from 1646 to 66600
We see that the data spans from January 1956 to August 1995 and it is a monthly time-series data.
The data follows a time order and there are no missing values
Periodicity of the dataset is monthly.

```{r}
ts.plot(ts.data, xlab = "Year", ylab = "Gas Production", main = "Australian Monthly Gas Production")
```

The gradual increase in production suggests the presence of trend component
The stable intra-year fluctuations indidcate the presence of seasonal component in time-series data.
The data clearly shows that production was stationary with stable seasonal fluctuations from 1956 to 1969. So, the overall series in non-stationary.
There onwards production has increased greatly showing strong presence of trend component.

```{r}
boxplot(ts.data ~ cycle(data), xlab = "Months", ylab = "Production", main = "Boxplot of Monthly Gas Production")
```

Production appears to be highest in the months of July and August and lowest in December-January

```{r}
monthplot(ts.data)
```

The vertical lines represting monthly production suggest highest production in July and also highest average production in July as represented by the horizontal lines. 

```{r}
ggseasonplot(ts.data, year.labels = TRUE, year.labels.left = TRUE)

ggseasonplot(ts.data, polar = TRUE)
```


There appears to be a sudden dip in production in March 1993 and a few other instances but the general seasonality suggests highest productoin in July or August
Production remained nearly constant till 1969 thereafter increased every year.
The above plots show that Production has seasonal fluctuations along with a trend. Thus there is evidence of multiplicative seasonality.

Decompose the data

```{r}
decomp.data.add = decompose(ts.data, type = "additive")
plot(decomp.data.add)

decomp.data.multi = decompose(ts.data, type = "multiplicative")
plot(decomp.data.multi)
```

No major changes between the multiplicative and additive models except for the random component.

```{r}
seasonal.indc = round(t(decomp.data.multi$figure),2)
seasonal.indc
```

By observing he seasonal indices we identify the highest production month as July(bearing the highest value of the seasonal component)
and lowest production month as January having the lowest seasonality component.


```{r}
stl.data = stl(log10(ts.data), s.window = "p")
plot(stl.data)

stl.data.3 = stl(ts.data, s.window = 3)
plot(stl.data.3)

stl.data$time.series[1:12,1]
data.season = exp(stl.data$time.series[1:12,1])
plot(data.season, type="l")
```

We observe constant seasonality in the data
Trend is highly significant as indicated by the small grey bars on the right and increases steadily from 1970 and then again attains normality 1990 onwards

Deseasonalize the data

```{r}
deseason.data = exp(stl.data$time.series[,2]) + exp(stl.data$time.series[,3])
ts.plot(deseason.data)
ts.plot(ts.data, deseason.data, col=c("red", "blue"), main="Comparison of Production and Deseasonalized Production")

```

Splitiing the time-series data in train and test data

```{r}
ts.train = window(ts.data, start = c(1956,1), end = c(1985,12), frequency = 12)

ts.test = window(ts.data, start = c(1986,1), frequency = 12)

autoplot(ts.train, series="Train") +
  autolayer(ts.test, series="Test") +
  ggtitle("Gas Production Traning and Test data") +
  xlab("Year") + ylab("Production") +
  guides(colour=guide_legend(title="Forecast"))
```


Random Walk with Drift

```{r}
ts.decomp.train = stl(log10(ts.train), s.window = "p")
ts.train.stl = forecast(ts.decomp.train, method = "rwdrift", h = 120)
plot(ts.train.stl)

vec1 = cbind(log10(ts.test),as.data.frame(forecast(ts.decomp.train, method = "rwdrift", h = 116))[,1])

vec = cbind(log10(ts.test),ts.train.stl$mean)

ts.plot(vec1, col = c("blue", "red"), main = "Australian Gas Production: Actual vs Forecast")
```

We notice that the forecasted data matches closely with the actual data from about 1986 to 1988 suggesting that the model can forecast accurately for 2 about years.   

```{r}
RMSE = round(sqrt(sum(((vec1[,1]-vec1[,2])^2)/length(vec1[,1]))),4)
RMSE

MAPE = round(mean(abs(vec1[,1]-vec1[,2])/vec1[,1]),4)
MAPE
```

Box-Ljung test:
H0: Residuals are Independent
Ha: Residuals are not Independent

```{r}
Box.test(ts.train.stl$residuals, type="Ljung-Box")
```
Accuracy measures give us a Root Mean Square Error value of 0.1483 and a Mean Absolute Percentage Error value of 0.0252. Also the Ljung-Box test gives us a very small p-value of 0.0008378 whcih signifies that the residuals are not independant.
This suggests that Random Walk with drift can give us a strong model which can forecast accurately for about 2 years the Australian Gas Production based on the given dataset.


ARIMA

Auto Regressive Induced Moving Average(ARIMA) requires a stationary time series so we check the data for stationarity using ADF test.

Augmented Dicky Fuller Test


𝐻0: Time series is non-stationary
𝐻1: Time series is stationary

```{r}
adf.test(ts.train)

# We check for stationarity after log transfor

ts.train.log = log10(ts.train)

adf.test(ts.train.log)
```
We observe that in both the cases the P-value after conducting Augmented Dicky Fuller test is much higher than 5% which leads us to conclude that the series in non-stationary.

```{r}
acf(ts.train, lag = 116, main = "ACF Time Series")
pacf(ts.train, lag = 116, main = "PACF Time Series")

acf(ts.train.log, lag = 116, main = "ACF Time Series")
pacf(ts.train.log, lag = 116, main = "PACF Time Series")

```
We try differencing the data to make the series stationary.

```{r}
ts.train.df = diff(ts.train)
plot(ts.train.df)

acf(ts.train.df, lag = 116)
pacf(ts.train.df, lag = 116)

adf.test(ts.train.df)

```

Autocorrelations are significant upto a very high lag value in the ACF plot. But after differencing the values change and we take q = 0.
The Partial Autocorrelations seem to suggest that about 1 past observation are significant hence, we take the p value for ARIMA as 1.

```{r}
arima.train = Arima(ts.train.df,c(1,1,0))
arima.train

hist(arima.train$residuals, col = "beige")

arima.fit = fitted(arima.train)

ts.plot(ts.train.df, arima.fit, col = c("red","blue"))
acf(arima.train$residuals)   


Box.test(arima.train$residuals, type="Ljung-Box")
```
The Ljung-Box test gives us a p-value of 60% whcih signifies that the residuals are independant.

```{r}

autoarima.train = auto.arima(ts.train.log, seasonal = TRUE)
autoarima.train

```

According to the auto arima result we should use the ARIMA(1,0,1)(0,1,2)[12] model for this time series data.
The lower AIC and BIC values indicate that the auto.arima model is better in terms of goodness of fit.

```{r}
tsf = Arima(ts.train.log, order = c(0,1,1), seasonal = c(0,1,2), method ="ML")
tsf
```

By using a seasonal ARIMA model with parameters obtained from auto.arima we get a better model.

```{r}
Box.test(tsf$residuals, type="Ljung-Box")
```

The Ljung-Box test gives us a p-value of over 80% whcih signifies that the residuals are stationary.

```{r}
ProdForecast = forecast(tsf, h = 116)
plot(ProdForecast)

vec2 = cbind(log10(ts.test),as.data.frame(forecast(tsf, h = 116))[,1])

ts.plot(vec2, col = c("blue", "red"), main = "Australian GAs Production: Actual vsForecast")

RMSE1 = round(sqrt(sum(((vec2[,1]-vec2[,2])^2)/length(vec2[,1]))),4)
RMSE1

MAPE1 = round(mean(abs(vec2[,1]-vec2[,2])/vec2[,1]),4)
MAPE1

```

The SARIMA model gives us a Root Mean Square error value of 0.0877 and Mean Absolute Percentage Error value of 0.0145 which suggests that this is a better model than RWDrift.

We proceed to forecast for the next 12 months using the SARIMA model.

```{r}
ts.final.arima = Arima(log10(ts.data), order = c(0,1,1), seasonal = c(0,1,2), method = 'ML')
ts.final.arima


ts.final.forecast = forecast(ts.final.arima , h = 12)
plot(ts.final.forecast)

```


